---
title: "DAS_Group_19"
output:
  pdf_document:
          latex_engine: xelatex
          number_sections: yes
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Introduction

The question for Group 19 is: Which factors influence the number of days an animal spends in the shelter before their final outcome is decided


```{r libraries,echo=FALSE,warning=FALSE,message=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(skimr)
library(moderndive)
library(sjPlot)
library(stats)
library(jtools)
library(readr)
library(patchwork)
```

```{r data processing,echo=FALSE,warning=FALSE,message=FALSE}
animals<-read.csv("C:/Users/Yuchen/Documents/tmp/DAS-Group-19/dataset19.csv")
```

We start by viewing the whole data at first.

```{r glimpse, echo=TRUE,eval=TRUE, include = TRUE}
glimpse(animals)
```
# Exploratory Data Analysis

We need to choose which GLM should be used for the model by plotting the densities of manufacture (dependent variable y)

We first investigate what kind of density function y (time in the shelter) obeys

The overall density function of time at shelter and the density functions for different animals are plotted below.

```{r densityy, echo=TRUE,eval=TRUE, include = TRUE}
p1<-ggplot(animals, aes(x = time_at_shelter))+   
  geom_density(color = "black", fill = "gray")+
  labs(x = "time stay in shelter", y = "density",
       title = "overall density function of time at shelter")

p2<-ggplot(animals, aes(x = time_at_shelter))+ 
  geom_density(aes(color = animal_type))+
  labs(x = "time stay in shelter", y = "density",
       title = "density functions for different animals")

p1+p2+plot_layout(ncol = 1)
```

As we can see, these two graphs have shown a typical poisson distribution which give us the direction of using the \textbf{logit link function} combined with the assumption that the response y (time at shelter) obeys the \textbf{poisson distribution} for \textbf{GLM} method later on.

But we will still observe the distribution of each variable (via boxplots and histograms)

Observing the time at shelter for different kinds of animals.

```{r animals, echo=TRUE,eval=TRUE, include = TRUE}
p11<-ggplot(data = animals, mapping = aes(x = factor(animal_type), y = time_at_shelter)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "kind of animals", y = "time stay in shelter",
       title = "different animals in shelter")  

p12<-ggplot(data = animals, mapping = aes(x = factor(animal_type), y = time_at_shelter, fill = animal_type)) +
  geom_col(position = "dodge") +
  labs(x = "kind of animals", y = "time stay in shelter",
       title = "different animals in shelter") 
p11+p12
```

Observing the time that animals stay in shelter in different month.

```{r month, echo=TRUE,eval=TRUE, include = TRUE}
p21<-ggplot(data = animals, mapping = aes(x = factor(month), y = time_at_shelter, fill = month)) +
  geom_col(position = "dodge") +
  labs(x = "month", y = "time stay in shelter",
       title = "the summary of months in time in shelter") 

p22<-ggplot(data = animals, mapping = aes(x = factor(month), y = time_at_shelter)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "month", y = "time stay in shelter",
       title = "summary of month in time in shelter")
p21+p22
```

After visualizing the data, we fond that there was no obvious changes but slightly different between the \textbf{first half year} and the \textbf{second half year} so we may classify time into two parts. (The month from 1 to 6 will be labelled as first half year and the month from 7 to 12 will be labelled as second half year)

```{r time, echo=TRUE,eval=TRUE, include = TRUE}
animals_used<-animals%>%
  mutate(time = ifelse(month > 6, "second half year", "first half year")) 
```

Observing the time that animals stay in shelter related to intake type.

```{r intake, echo=TRUE,eval=TRUE, include = TRUE}
p31<-ggplot(data = animals, mapping = aes(x = factor(intake_type), y = time_at_shelter, fill = intake_type)) +
  geom_col(position = "dodge") +
  labs(x = "intake_type", y = "time stay in shelter",
       title = "the summary of intake_type in time in shelter") 

p32<-ggplot(data = animals, mapping = aes(x = factor(intake_type), y = time_at_shelter)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "intake_type", y = "time stay in shelter",
       title = "summary of intake_type in time in shelter") 
p31+p32+plot_layout(ncol = 1)
```

Observing the time that animals stay in shelter related to outcome type.

```{r outcome, echo=TRUE,eval=TRUE, include = TRUE}
p41<-ggplot(data = animals, mapping = aes(x = factor(outcome_type), y = time_at_shelter, fill = month)) +
  geom_col(position = "dodge") +
  labs(x = "outcome_type", y = "time stay in shelter",
       title = "the summary of outcome_type in time in shelter") 
p42<-ggplot(data = animals, mapping = aes(x = factor(outcome_type), y = time_at_shelter)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "outcome_type", y = "time stay in shelter",
       title = "summary of outcome_type in time in shelter") 
p41+p42+plot_layout(ncol = 1)
```

Observing the time that animals stay in shelter related to chip status.

```{r chip, echo=TRUE,eval=TRUE, include = TRUE}
p51<-ggplot(data = animals, mapping = aes(x = factor(chip_status), y = time_at_shelter, fill = chip_status)) +
  geom_col(position = "dodge") +
  labs(x = "chip_status", y = "time stay in shelter",
       title = "the summary of chip_status in time in shelter") 
p52<-ggplot(data = animals, mapping = aes(x = factor(chip_status), y = time_at_shelter)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "chip_status", y = "time stay in shelter",
       title = "summary of ochip_status in time in shelter")
p51+p52+plot_layout(ncol = 1)
```

# Model Analysis

After observing all the variables, we fond that there was no obvious linear relationship. But as we mentioned before, we can use \textbf{GLM} method with family=poisson(link="log") for fitting the data. And the formula we will use is given by: 
$$y_{i}$$ $$\sim$$ $$Poisson\left(\lambda_{i}\right)$$
                      $${log}\left(\lambda_{i}\right)=\beta_{0}+\beta_{1}x_{1i}+\cdots+\beta_{j}x_{ji}$$
Where $$\lambda_{i}=exp\left(\beta_{0}+\beta_{1}x_{1i}+\cdots+\beta_{j}x_{ji}\right)$$
and $$\frac{\lambda_{ji}}{\lambda_{0}}=exp\left(\beta_{j}\right)$$
where $\lambda_{ji}$ is the value of $\lambda_{i}$ when only $x_{ji}$ is 1 and $\lambda_{0}$ is the exp of $\beta_{0}$.

Since all the explanatory variables are characteristic, the exp of $\beta_{j}$ means that the mean of $y_{i}$ will be $e^{\beta_{j}}$ times of the mean where the $j_{th}$ factor changes from 0 to 1.

As a result, if the exp of coefficient is closer to 1, the more unlikely the factor will affect the response term.
                    
We first try the full model for fitting the whole data.

```{r model1, echo=TRUE,eval=TRUE, include = TRUE}
animals_used<-animals_used%>%
  select(-c("month"))
animals_used$year<-as.character(animals_used$year)
mod.loglinear <- glm(time_at_shelter ~ year + animal_type + intake_type + outcome_type + chip_status + time, data = animals_used, family = poisson(link = "log"))
summary(mod.loglinear)
AIC(mod.loglinear)
```
We fond that collinearity exists between different animal types. We then check the data grouped by animal type and fond that there were only 2 rows of data for \textbf{BIRD}. Since we know that small size of data would cause collinearity, we decided not to include the data for BIRD.

```{r Bird, echo=TRUE,eval=TRUE, include = TRUE}
animals_used<-animals_used%>%
  filter(animal_type != "BIRD")
```

Then we use the new version of data for building the model.

```{r finalmodel, echo=TRUE,eval=TRUE, include = TRUE}
mod.loglinear_alt <- glm(time_at_shelter ~ year + animal_type + intake_type + outcome_type + chip_status + time, data = animals_used, family = poisson(link = "log"))
summary(mod.loglinear_alt)
AIC(mod.loglinear_alt)
```

We then have a new model where no p values for explanatory variables are greater than 0.05.

We can now have a look at the exp of coefficient plotted and listed below:
```{r plotmodel, echo=TRUE,eval=TRUE, include = TRUE}
confint(mod.loglinear_alt)
plot_model(mod.loglinear_alt, show.values = TRUE, title = "exp-mean", show.p = FALSE,vline.color = "darkblue")
```

```{r exp, echo=TRUE,eval=TRUE, include = TRUE}
mod.loglinear_alt %>%
   coef() %>%
   exp()
```

# Conclusion

We then conclude that the factors: intake type and outcome type affect the time at shelter the most.

Residuals can be seen as follow the assumption. Actually, the residuals should follow the assumptions theoretically by the theory of GLM.

```{r residuals, echo=TRUE,eval=TRUE, include = TRUE}
animals_used <- animals_used %>%
  mutate(counts_pred = fitted(mod.loglinear_alt))
p1<-ggplot(mod.loglinear_alt, aes(x = mod.loglinear_alt$residuals))    
p1 + geom_boxplot(color = "black", fill = "gray")
```


